---
title: Credit Suisse Twitter Sentiment
author: 'Elio Bolliger'
date: '2023-03-21'
summary: 'In this short project, we analyse tweets from Twitter that contain the keyword "Credit Suisse". We then use a Natural Language Processing model to predict the sentiment of those tweets.'
slug: []
categories:
tags:
  - Financial Markets
  - Credit Suisse
  - NLP
links:
- icon: twitter
  icon_pack: fab
  name: Follow
  url: https://twitter.com/BolligerElio
slides: ""
math: true
description: Description for the page
bibliography: [bibliography.bib]
editor_options: 
  chunk_output_type: console
---

<style>
  .bottom-three {
     margin-bottom: 12cm;
  }
</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, fig.height = 4, warning = FALSE, out.width = "100%"
)

if (!require("ggplot2")) {install.packages("ggplot2"); library('ggplot2')}
if (!require("tidyverse")) {install.packages("tidyverse"); library('tidyverse')}
if (!require("matrixStats")) {install.packages("matrixStats"); library('matrixStats')}
if (!require("tidymodels")) {install.packages("tidymodels"); library('tidymodels')}
if (!require("lme4")) {install.packages("lme4"); library('lme4')}
if (!require("haven")) {install.packages("haven"); library('haven')}
if (!require("miceadds")) {install.packages("miceadds"); library('miceadds')}
if (!require("rmarkdown")) {install.packages("rmarkdown"); library('rmarkdown')}
if (!require("reshape2")) {install.packages("reshape2"); library('reshape2')}
if (!require("plotly")) {install.packages("plotly"); library('plotly')}
if (!require("knitr")) {install.packages("knitr"); library('knitr')}
if (!require("kableExtra")) {install.packages("kableExtra"); library('kableExtra')}
if (!require("htmltools")) {install.packages("htmltools"); library('htmltools')}
if (!require("data.table")) {install.packages("data.table"); library('data.table')}
if (!require("zoo")) {install.packages("zoo"); library('zoo')}
if (!require("processx")) install.packages("processx")

theme_set(theme_bw())
```


```{r}
# Move this css tag outside the chunk to control the width of text
# on the page.
# <style type="text/css">
# .main-container {
#   max-width: 1000px;
#   margin-left: auto;
#   margin-right: auto;
# }
# </style>

```



```{r, include = FALSE}

# load in separately downloaded data:

dat <- fread("/Users/ebollige/Dropbox/3_PhD/Projects/TWITTER_CREDITSUISSE/sentimentdf/tweets_1.csv")

dat1 <- dat %>%
  dplyr::select(date,sentiment) %>%
  dplyr::mutate(sn = case_when(sentiment == "negative" ~ -1,
                                      sentiment == "positive" ~ 1,
                                      sentiment == "neutral" ~ 0)) %>%
  dplyr::mutate(date = as.Date(date)) %>%
  group_by(date) %>%
  summarise(sn = mean(sn,na.rm = TRUE),
            n = n())

dat <- fread("/Users/ebollige/Dropbox/3_PhD/Projects/TWITTER_CREDITSUISSE/sentimentdf/tweets_2.csv")

dat2 <- dat %>%
  dplyr::select(date,sentiment) %>%
  dplyr::mutate(sn = case_when(sentiment == "negative" ~ -1,
                                      sentiment == "positive" ~ 1,
                                      sentiment == "neutral" ~ 0)) %>%
  dplyr::mutate(date = as.Date(date)) %>%
  group_by(date) %>%
  summarise(snupdate = mean(sn,na.rm = TRUE),
            nupdate = n())


dat <- fread("/Users/ebollige/Dropbox/3_PhD/Projects/TWITTER_CREDITSUISSE/sentimentdf/tweets_3.csv")

dat3 <- dat %>%
  dplyr::select(date,sentiment) %>%
  dplyr::mutate(sn = case_when(sentiment == "negative" ~ -1,
                                      sentiment == "positive" ~ 1,
                                      sentiment == "neutral" ~ 0)) %>%
  dplyr::mutate(date = as.Date(date)) %>%
  group_by(date) %>%
  summarise(sn = mean(sn,na.rm = TRUE),
            n = n())

# fill gaps october 12 to october 19:
dat <- fread("/Users/ebollige/Dropbox/3_PhD/Projects/TWITTER_CREDITSUISSE/sentimentdf/tweets_4.csv")

dat4 <- dat %>%
  dplyr::select(date,sentiment) %>%
  dplyr::mutate(sn = case_when(sentiment == "negative" ~ -1,
                                      sentiment == "positive" ~ 1,
                                      sentiment == "neutral" ~ 0)) %>%
  dplyr::mutate(date = as.Date(date)) %>%
  group_by(date) %>%
  summarise(sn = mean(sn,na.rm = TRUE),
            n = n()) %>%
  dplyr::filter(date >= "2022-10-12" & date <= "2022-10-19")



```


```{r, include = FALSE}
# update data
dat <- rbind(dat1,dat3, dat4)

# merge data:
datmerge <- merge(dat,dat2, by=c("date"), all = TRUE)
datmerge <- datmerge %>%
  dplyr::mutate(sn = ifelse(is.na(snupdate)== F, snupdate, sn )) %>%
  dplyr::mutate(n = ifelse(is.na(nupdate)== F, nupdate, n )) %>%
  dplyr::select(date, sn, n)


# add data about stock market prices:
cs <- fread("/Users/ebollige/Dropbox/3_PhD/Projects/TWITTER_CREDITSUISSE/CSGN_shareprice.csv") %>%
  dplyr::select(Date, Close) %>%
  dplyr::rename(date = Date) %>%
  dplyr::mutate(date = as.Date(date)) %>%
  arrange(date) %>%
  dplyr::mutate(gr = (Close - lag(Close))/lag(Close)*100) %>%
  dplyr::filter(date >= "2021-01-01")

# merge data:
fdat <- merge(datmerge,cs, all.y = TRUE, all.x = TRUE)

# standardise the sentiment measure and linearly interpolate data for growth rate of stocks
fdat <-  fdat %>%
  ungroup() %>%
  dplyr::mutate(sumobs = sum(n, na.rm = TRUE)) %>%
  dplyr::mutate(wm = (sn * n)/sumobs ) %>%
  dplyr::mutate(resmeandat= ifelse(date <= "2023-03-01", wm, NA)) %>%
  dplyr::mutate(means = mean(resmeandat,na.rm = TRUE)) %>%
  dplyr::mutate(sds = sd(resmeandat,na.rm = TRUE) ) %>%
  dplyr::mutate(sentstandard = (wm-means)/sds ) %>%
  ungroup() %>%
  arrange(date)  %>%
    mutate(time=seq(1,n())) %>%
      mutate(grip=approx(time,gr,time)$y) %>%
      select(-time)

correlation <- round(cor(fdat$grip,fdat$sentstandard, use="complete.obs"),2)
correlation

```





```{r, out.width="100%"}
fdat <- fdat %>%
  dplyr::filter(date >= "2022-03-01")


sum <- sum(fdat$n,na.rm = TRUE)

fig <- plot_ly()
# Add traces
fig <- fig %>% add_trace(x = ~fdat$date, y = ~fdat$grip, name = "Credit Suisse Stock Price", mode = "lines+markers", 
                         type = "scatter", 
                         marker=list(symbol = 'x',
                                      size = 3,
                                      opacity = 1, line = 
                                        list(width = 1, color = "steelblue")),
                          line = list(color = "steelblue")
                         )

ay <- list(
  tickfont = list(color = "firebrick"),
  overlaying = "y",
  side = "right",
  title = "<b>Sentiment</b>")

fig <- fig %>% add_trace(x = ~fdat$date, y = ~fdat$sentstandard, name = "Sentiment on Twitter", yaxis = "y2", mode = "lines+markers", type = "scatter", marker=list(color = "firebrick", symbol = 'triangle-up',
                                      size = 3,
                                      opacity = 1, line = 
                                        list(width = 1, color = "firebrick")),
        line = list(color = "firebrick"))

# Set figure title, x and y-axes titles
fig <- fig %>% layout(
  title = "", yaxis2 = ay,
  xaxis = list(title="Date"),
  yaxis = list(title="<b>Stock Price, Daily Change (%)</b>"),
        legend = list(x = 0.1, y = 0.9)
)%>%
  layout(plot_bgcolor='white', 
         xaxis = list( 
           zerolinecolor = '#ffff', 
           zerolinewidth = 2), 
          yaxis = list(
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            range=c(-25,25)
          ),
        yaxis2 = list(
            range=c(-5,5)
            #range=c(-0.01,0.01)
        ),
        width = 1000, height = 800, autosize = F
        ) 

fig <- fig %>% add_annotations(x=as.Date("2022-09-22"), y=-12,
            text="Rumours about capital hike",
            showarrow=T,
            arrowhead=1,
            ax=-100)

fig <- fig %>% add_annotations(x=as.Date("2022-10-02"), y=-20,
            text="New rumours about capital hike",
            showarrow=T,
            arrowhead=1, ax = -100)

fig <- fig %>% add_annotations(x=as.Date("2022-10-27"), y=-24,
            text="Confirmation capital raise and strategy release",
            showarrow=T,
            arrowhead=1, ax = 100)

fig <- fig %>% add_annotations(x=as.Date("2022-11-01"), y=-12.3,
            text="S&P donwgraded CS",
            showarrow=T,
            arrowhead=1, ax = 100)

fig <- fig %>% add_annotations(x=as.Date("2023-03-15"), y=-24,
            text="Saudi NB announcement",
            showarrow=T,
            arrowhead=1, ax = -100, ay = -60)

fig


```

<p class="bottom-three">
     
  The graph above displays the **sentiment** predicted for `r format(sum, big.mark=",")` tweets that contain the **keyword "Credit Suisse"**, along with the **daily stock price change** (in %) of Credit Suisse shares. To predict the sentiment of a tweet, I use a Natural Language Processing model trained on financial texts (FinBERT, @araci2019finbert).
  
  
  In the plot, the sentiment measure represents the **average sentiment per day** of all collected tweets. The sentiment is then weighted by the number of tweets collected that day, divided by the total number of tweets. The measure is normalized (mean subtracted and divided by its standard deviation).^[The mean and standard deviation are calculated without observations of March 2023.] Higher values correspond to a more positive sentiment, whereas lower values represent a negative sentiment. For the daily change in stock prices, missing values over the weekend are linearly interpolated to improve the readability of the chart. The **correlation** between the two series is `r correlation`.
  

A few notes on the **negative spikes in sentiment** values in the graph and **news media reporting** around the same time: 

- **On September 22nd, 2022**, **news emerged** that Credit Suisse **might** consider splitting their investment bank (@reuters1) and that they were sounding out investors about a **capital hike** (@reuters2).

- **Around October 2nd, 2022**, news emerged that Credit Suisse **might** be looking to **raise capital**. At the same time, the price of Credit Default Swaps increased strongly (@cnbc1).

- **Around October 27th, 2022**, Credit Suisse **confirmed raising capital**, announced a job cut, and informed about their new strategy (@cs1, @reuters3).

- **On November 1st, 2022**, **S&P downgraded Credit Suisse's credit rating** to BBB- (@wsj1).

- **On March 15th, 2023**, the chairman of **Saudi National Bank** announced that the bank **wouldn't boost its share of the bank** (@bloomberg).


For more information, there is a growing literature that investigates the links between social media on financial markets. See, for example, @goutte2022actions, @broadstock2019social, @antweiler2004all, @sprenger2014tweets.

</p>

<br>

**References and Sources**
